---
title: "RTSA HW5 Q2 - Varshini Yanamandra"
output: html_notebook
---

```{r}
# Installing required libraries
library(faraway)
library(MASS)

# Loading data
myData = faraway::sat
head(myData)
```
```{r}
dim(myData)
```

Part (a)
```{r}
# Creating the multiple linear regression model
myModel <- lm(total ~ expend + salary + ratio + takers, data = myData)
summary(myModel)
```

Part (b)
```{r}
hatv = hatvalues(myModel)
head(sort(hatv, decreasing = T))
sum(hatv)
```
There are 50 rows (n = 50) and 4 regressors (p = 4)
Hence, a point is influential if its leverage is > 2*(p+1)/n = 0.2
There are 4 states with leverage value > 0.2
```{r}
# Half normal plot
halfnorm(hatv, nlab = 4, labs = state.name, ylab = "Hat Value")
```

Part (c)
```{r}
# jackknife residuals or studentised residuals
#jkres <- studres(myModel)
#head(jkres[order(abs(jkres), decreasing = T)])
stud <- rstudent(myModel)
jkres <- stud*(44/(45-stud^2))^0.5
head(jkres[order(abs(stud), decreasing = T)])
```

```{r}
# Bonferroni correction
qt(0.025/50, 44)
```
The largest outlier, West Virginia, is not significant at the 5% level as -3.491336 is greater than -3.525801.

Part (d)
```{r}
cook <- cooks.distance(myModel)
head(sort(cook, decreasing = T))
halfnorm(cook, 3, labs=state.name, ylab = "Cook's Distances")
```
The Cook's distance for Utah is much greater than the state with the next highest Cook's distance (West Virginia). However, since none of the states have a Cook's distance > 1, they can be considered non-influential.

Part (e)
```{r}
vif_values <- vif(myModel)
vif_values
mean(vif_values)
```
While the greatest VIP (expend) is not > 10, the mean VIF is 5.717713, which is much greater than 5. This suggests a good chance of multicollinearity.