---
title: "RTSA HW 7"
output: html_notebook
author: "Varshini Yanamandra"
---

QUESTION 3
The answers to part (c) - the observations - are written in the bottom of each graph.

```{r}
# part a
set.seed(12345)
y1 = arima.sim(model = list(ar=0.6), n = 100)
par(mar=c(5,4,4,2))
acf(y1, main = 'Autocorrelations', lag.max = 99, ci.col = 'black')
acf(y1, main = 'Autocorrelations', lag.max = 10, ci.col = 'black')
```
```{r}
# part b
set.seed(12345)
y2 = arima.sim(model = list(ar=-0.6), n = 100)
par(mar=c(5,4,4,2))
acf(y2, main = 'Autocorrelations', lag.max = 99, ci.col = 'black')
acf(y2, main = 'Autocorrelations', lag.max = 30, ci.col = 'black')
```
```{r}
# part c
set.seed(12345)
y3 = arima.sim(model = list(ar=0.95), n = 100)
par(mar=c(5,4,4,2))
acf(y3, main = 'Autocorrelations', lag.max = 99, ci.col = 'black')
acf(y3, main = 'Autocorrelations', lag.max = 70, ci.col = 'black')
```
```{r}
# part d
set.seed(12345)
y4 = arima.sim(model = list(ar=0.3), n = 100)
par(mar=c(5,4,4,2))
acf(y4, main = 'Autocorrelations', lag.max = 99, ci.col = 'black')
acf(y4, main = 'Autocorrelations', lag.max = 10, ci.col = 'black')
```
Observations:
The lag until when the ACF is significantly distant from 0 increases as the absolute value of phi increases and decreases as the absolute value of phi decreases, as seen from these four examples.

QUESTION 4

The stationary processes are the processes in parts (a) and (b).
```{r}
# part a
set.seed(12345)
y1 = arima.sim(model = list(ar=c(-0.2, 0.48)), n = 100)
par(mar=c(5,4,4,2))
acf(y1, main = 'Autocorrelations', lag.max = 30, ci.col = 'black')
pacf(y1, main = "Partial ACFs", lag.max = 30, ci.col = 'black')
```
```{r}
# part b
set.seed(12345)
y2 = arima.sim(model = list(ar=c(-0.6), ma=c(1.2)), n = 100)
par(mar=c(5,4,4,2))
acf(y2, main = 'Autocorrelations', lag.max = 30, ci.col = 'black')
pacf(y2, main = "Partial ACFs", lag.max = 30, ci.col = 'black')
```

