---
title: "RTSA HW 8"
output: html_notebook
author: "Varshini Yanamandra"
---
Question 3

```{r}
# loading packages for question 3
library(ggfortify)
```
```{r}
data("AirPassengers")
myData <- AirPassengers
class(myData) # confirming it is a time-series dataset
myData
```
```{r}
# time series plot
ts.plot(myData) # way 1
autoplot(myData) # way 2
```
```{r}
# acf plot
par(mar=c(5,4,4,2))
acf(myData, main = "Autocorrelation", lag.max = 160)
acf(myData)
```
```{r}
# fitting an ARMA(1, 1) model to the data
ARMA_model <- arima(myData, order = c(1, 0, 1))
print(ARMA_model)
ts.plot(myData)
ARMA_fit <- myData - resid(ARMA_model)
points(ARMA_fit, type = "l", col = 2, lty = 2)
```
The model has been fitted very well with the fitted values being very close to the actual values. However, there is a forward lag in the fitted values, as in, the trend (graph) of the fitted values is leading the trend of the actual data by a small value.

Question 4

```{r}
data("sunspots")
sunData <- sunspots
class(sunData) # confirming it is a time-series dataset
```
```{r}
# time series plot
ts.plot(sunData)
```
```{r}
# acf plot
par(mar=c(5,4,4,2))
acf(sunData, main = "Autocorrelation", lag.max = 10000)
acf(sunData)
```
```{r}
# fitting an ARIMA(0, 1, 1) model to the data
IMA_model <- arima(sunData, order = c(0, 1, 1))
print(IMA_model)
ts.plot(myData)
IMA_fit <- myData - resid(IMA_model)
points(IMA_fit, type = "l", col = 2, lty = 2)
```
```{r}
# fitting an ARIMA(1, 1, 0) model to the data
ARI_model <- arima(sunData, order = c(1, 1, 0))
print(ARI_model)
ts.plot(myData)
ARI_fit <- myData - resid(ARI_model)
points(ARI_fit, type = "l", col = 2, lty = 2)
```
```{r}
# AIC of the two models - the same as what is given in the summary tables for the models
ima_aic = AIC(IMA_model)
ari_aic = AIC(ARI_model)
ima_aic
ari_aic
```
As the ARIMA(0,1,1) model has a lower AIC, it would be better to pick that model based on AIC alone.
```{r}
# Ljung-Box test
Box.test(IMA_model$residuals, lag = 12 ,type = "Ljung", fitdf = 2)
```
The p-value is much smaller than 0.05 (5% significance), so we can say that the residuals are not independent (and are correlated) by rejecting the null hypothesis.
At 5% significance, the null hypothesis (that the residuals are not correlated) is rejected, meaning at least of the correlations is non-zero for the residuals.
